Choosing climate forcing for numerical simulations of the Cordilleran ice sheet

### 2013-02-17 sections/model.tex
Included text from the old draft as comments
Added subsection 3.1 Surface mass-balance

\begin{figure}\begin{framed}
	\textbf{Technical note for Constantine}
	\todo{Remove this note}

@@

>>>>

\begin{figure}\begin{framed}
	\textbf{Technical note for Constantine}

@@

\subsection{Surface mass-balance}

\begin{table}[t]
	\caption{Values of the positive degree-day model parameters.}
	\label{tab:pdd}
	\vskip4mm
	\centering
	\begin{tabular}{ll}
		\tophline
		Parameter& Value\\
		\middlehline
		degree-day factor for snow & 3\,\unit{mm\,K^{-1}\,day^{-1}}\\
		degree-day factor for ice & 8\,\unit{mm\,K^{-1}\,day^{-1}}\\
		standard deviation of temperatures & 5\,K\\
		refreezing faction & 0.6 \\
		\bottomhline
	\end{tabular}
\end{table}

Monthly temperatures and precipitation rates are converted to surface accumulation and ablation using a temperature-index (positive degree-day) model\needref. Accumulation is computed as the amount of precipitation that falls at times when temperature is below freezing \todo{Check PISM doc}. Ablation is a transfer function of the number of positive degree-day, defined as the sum of mean daily temperatures through the year. Positive degree-day are computed by the \citet{calov-greve-2011} integrand to account for day-to-day variability which is not represented in the monthly climate forcing. The ablation model use a different coefficient for snow and ice and accounts for refreezing. Model parameters are explicited in table~\ref{tab:pdd}

Although a positive degree-day model is a very crude approximation of processes occurring at the glacier surface, we will show that it captures reasonable regional patterns of accumulation and ablation when using accurate climate data.

\todo[inline]{
	\textbf{Include the \citet{calov-greve-2011} integrand formula?}

	The number of positive degree-days, defined as the integrand of positive temperatures over time, is computed annually using a formula derived by \citet{calov-greve-2011} that accounts for day-to-day variability of temperature by assuming a normal distribution of variance $\sigma=5\degree C$ along a central (input) value $T$:

	\begin{equation}
		\mathrm{PDD} = \int_{t_0}^{t_1} dt \left[
			\frac{\sigma}{\sqrt{2\pi}}
			\exp\left({-\frac{T(t)^2}{2\sigma^2}}\right)
			+\frac{T(t)}{2}
			\mathrm{erfc} \left(-\frac{T(t)}{\sqrt{2}\sigma}\right)
		\right]
	\end{equation}

Where $[t_0; t_1]$ is one year and $\mathrm{erfc}$ is the complementary error function.}

%\subsection{Ice flow dynamics}
%To resolve ice dynamics, PISM specifically uses a heuristic combination of two classical and well-based zeroth-order shallow-ice approximations through a weighting function \citep{bueler-brown-2009}. The Shallow Shelf Approximation \citep[SSA,][]{morland-1987,weis-etal-1999} is used as a ``sliding law'' for the Shallow Ice Approximation \citep[SSA,][]{hutter-1983}. Depth-averaged velocities $\vec{v}$ are computed as a weighted sum of SIA velocities $\vec{v_{\mathrm{SIA}}}$ and SSA velocities $\vec{v_{\mathrm{SSA}}}$ by
%
%\begin{equation}
%	\vec{v} = f(|\vec{v_{\mathrm{SSA}}}|)\vec{v_{\mathrm{SIA}}}
%	+ (1-f(|\vec{v_{\mathrm{SSA}}}|))\vec{v_{\mathrm{SSA}}},
%\end{equation}
%
%with
%
%\begin{equation}
%	f(|\vec{v_{\mathrm{SSA}}}|) = 1
%	- \frac{2}{\pi}\arctan{\frac{|\vec{v_{\mathrm{SSA}}}|^2}{100^2}}.
%\end{equation}
%
%This approach, although using a heuristic, allows realistic representation of horizontal shear and longitudinal stresses where they matter \citep{bueler-brown-2009}.
%
%
%The SSA velocities are determined by a a pseudo-plastic sliding law in which the yield stress is depends on the availability of basal water and a prescibed till friction angle. We apply lower friction angles at lower elevation to model more slippery beds due to the presence of marine sediments.
%
%\begin{equation}
%	\phi = \left\{\begin{array}{llc}
%		10\,\unit{\degree C} & \mathrm{for} &               z<  0\,\unit{m} \\
%		z/10 + 10°           & \mathrm{for} &   0\,\unit{m}<z<200\,\unit{m} \\
%		30\,\unit{\degree C} & \mathrm{for} & 200\,\unit{m}<z               \\
%	\end{array}\right.
%\end{equation}
%
%\subsection{Glacier bed interface}
%
%The ice-sheet bed topography responds to the ice-loads following a regional isostasy model with flexure and relaxation.
%
%Geothermal heat flux is prescribed to 70\,\unit{W\,m^{-2}}
%
%\subsection{Simulation design and numerical implementation}
%
%The simulations were run with PISM, a parallel ice sheet model (version stable 0.5.11). Documentation on the code can be found in several publications \citep[e.g.,][]{bueler-brown-2009,aschwanden-etal-2012} and more extensively online \citep[{\url{http://www.pism-docs.org}},][]{web:pism}.

%Model equations are discretized horizontally on a 10~km-resolution, 150 by 300 points grid, and vertically within the ice on 21 irregularly-spaced layers. Time-stepping is adaptive, but a maximum allowed time step of 0.08 is used to ensure monthly capture of climate data by the surface mass-balance model.
%
%The simulations were run in parallel on 9~computation nodes of 8~cores each, for a total of 72~cores, on the supercomputer ekman at the PDC center for high performance computing of Kungliga Tekniska högskolan in Stockholm. Each simulation completed in about 10~hours.

%In this study we model glacial inception and growth of the Cordilleran ice sheet to a state nearing its last glacial maximum. For each present-day climate forcing we run a series of six simulations by applying a constant, homogeneous temperature offset of 4 to 9 \unit{\degree C} over the entire model domain. Each simulation initializes from ice-free conditions and is run for 10~000~years, a time interval we think to be representative of the rapid growth of the ice sheet from nearly ice-free conditions to last glacial maximum extent \needref. No precipitation corrections were used.
%
